---
title: "R Notebook"
output: html_notebook
---

For File Reading: crimedata.csv

```{r}
crimedata <- read.csv(file.choose())

```

For data Cleaning:

```{r}
vars<-c("state","population","householdsize","NumUnderPov","TotalPctDiv","PersPerFam","PctWorkMom","NumKindsDrugsSeiz","PctUsePubTrans","PolicAveOTWorked","LandArea","NumKidsBornNeverMar","PctLess9thGrade","PctNotHSGrad","PctBSorMore","racepctblack","racePctWhite","racePctAsian","racePctHisp","pctUrban","medFamInc","medIncome","pctWWage","PctUnemployed","MedRent","PopDens","nonViolPerPop","ViolentCrimesPerPop")
crimedata<-crimedata[,vars]
crimedata<-na.omit(crimedata)
```

For Model Testing:

```{r}
testmod<-lm(ViolentCrimesPerPop~.,data=crimedata)
summary(testmod)

testmod1<-lm(ViolentCrimesPerPop~.-state,data=crimedata)
summary(testmod1)
```
Finds relationship between variables as per the dependent variable
```{r}
lm1<-lm(ViolentCrimesPerPop~1,data=crimedata)
lmALL<-lm(ViolentCrimesPerPop~.-state,data=crimedata)
lmSTEP<-step(lm1,scope=formula(lmALL),direction="forward",trace=0)

summary(lmSTEP)
```

```{r}

summary(crimedata)

```

```{r}
lm1<-lm(medIncome~1,data=crimedata)
lmALL<-lm(medIncome~.-state,data=crimedata)
lmSTEP<-step(lm1,scope=formula(lmALL),direction="forward",trace=0)

summary(lmSTEP)
```

To Install Library

```{r}

install.packages("ggplot2")
```

Plot a relationship between the income in each state:

```{r}
library(ggplot2)
qplot(medIncome, state, data=crimedata, color = medIncome, size=10)

```
Plot a regression model between median income and non violent crimes


```{r}
library(ggplot2)
qplot(medIncome, nonViolPerPop, data=crimedata, color = nonViolPerPop, size=16)

```

Plot a regression model between median income and violent crime per population:

```{r}
library(ggplot2)
qplot(medIncome, ViolentCrimesPerPop, data=crimedata, color = ViolentCrimesPerPop , size=nonViolPerPop)


```
To Make the color of scattered data transparent:


```{r}
library(ggplot2)
qplot(medIncome, ViolentCrimesPerPop, data=crimedata, color = ViolentCrimesPerPop , size=nonViolPerPop, alpha=0.5)

```

```{r}

library(ggplot2)
qplot(log(medIncome), log(ViolentCrimesPerPop), data=crimedata, color = "red" , size=16) + facet_wrap(~ ViolentCrimesPerPop )

```

Draws linear regression line on the model:

```{r}
qplot(log(medIncome), log(ViolentCrimesPerPop), data = crimedata, color= ViolentCrimesPerPop, size = 12) +geom_smooth(method = "lm", size = 1 )
ViolentCrimesPerPop.lm

```
draw a scatterplot


```{r}
plot(log(medIncome)~ log(ViolentCrimesPerPop), data= crimedata, ylab= "violent crimes", xlab= "Median Income")
abline(ViolentCrimesPerPop.lm$coefficients)
```
Anova Table:


```{r}
anova(ViolentCrimesPerPop.lm)

```
Statistical values calculations:


```{r}
summary(crimedata.lm)
```

Fitting linear regression model:

```{r}
x_range <- range(crimedata$medIncome)
y_range <- range(crimedata$ViolentCrimesPerPop)
Range <- c(x_range, y_range)
crimedata.lm <- lm(ViolentCrimesPerPop ~ medIncome, data = crimedata)

summary(crimedata.lm)

par(pty = "s")
plot (log(medIncome) ~ ViolentCrimesPerPop, data = crimedata,
     xlab= "median Income",
     ylab= "Violent Crimes Per Pop",
     pch= 16,
     col= rgb(67, 110, 238, alpha = 130, max = 255))
print(Range)

abline(crimedata.lm$coeff)

```

```{r}
crimedata.lm <- lm(ViolentCrimesPerPop ~ medIncome, data=crimedata)
names(crimedata.lm)
```

```{r}
crimedata.lm
```

```{r}
summary(crimedata.lm)
```
PLotting scatterplot:

```{r}
plot (medIncome ~ ViolentCrimesPerPop, data= crimedata)
```
Plotting linear regression line:


```{r include = TRUE}
crimedata.lm <- lm(log(medIncome) ~ ViolentCrimesPerPop, data= crimedata)
plot (log(medIncome) ~ ViolentCrimesPerPop, data = crimedata, ylab = "Violent Crimes", xlab = "Median Income")
abline(crimedata.lm$coefficients, col=("red"))



```
Removing outliers:


```{r include = TRUE}
plot (log(medIncome) ~ ViolentCrimesPerPop, data = crimedata, ylab = "Median Income", xlab = "Violent Crimes", subset=-1)
abline(crimedata.lm$coefficients)
crimedataout.lm <- lm(log(medIncome) ~ ViolentCrimesPerPop,subset = -1, data = crimedata)
abline(crimedataout.lm$coefficients, col= "red")



```

Making assumptions on the model:

```{r include = TRUE}
crimedata.lm$coefficients
summary(crimedata.lm)

```
calculating confidence interval:

```{r}
confint(crimedata.lm)
```

Ploting anova table:

```{r}
anova(crimedata.lm )



```

importing residual file

```{r}
 
sresid <- rstandard(crimedata.lm)

```


ploting  residuals:

```{r}
par(pty = "s")
qqnorm(sresid, col= "royalblue")
qqline(sresid)
```

ploting standardised residual against violent crimes:


```{r}

plot(sresid ~ ViolentCrimesPerPop, data = crimedata)
abline(h = seq(-2,2), lty=2)
abline(h=0)

```

```{r}

plot(sresid ~ crimedata.lm$fitted, xlab="fitted values")
abline(h = seq(-2,2), lty=2)
abline(h=0)


```
verifing the model:

```{r}
plot(crimedata.lm)
```
calculating correlatioin:
```{r}
correlation <- cor(crimedata$ ViolentCrimesPerPo, crimedata$medIncome, use )
```

box plot 

```{r}
boxplot(crimedata$ViolentCrimesPerPo ~ crimedata$medIncome,
        xlab = "ViolentCrimesPerPo",
        ylab = "medIncome",
        main = "Box Plot of Income by Crime Rate")
```
box plot of all variables:


```{r}
vars <- names(crimedata)[-1]
boxplot_list <- list()
for (var in vars) {
  boxplot_list[[var]] <- boxplot(crimedata[[var]],
                                 main = paste("Box Plot of", var),
                                 xlab = var)}
par(mfrow = c(3, 3))
for (var in vars) {
  boxplot_list[[var]]
}
```
